---
title: "Cleaning_2014_data"
author: "Elise Boos"
date: '2022-08-15'
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages etc}
getwd()
#packages
require(tidyverse)
require(dplyr)
require(sf)
require(leaflet)
require(mapview)
require(readxl)
require(lubridate)

library(unmarked)
library(AICcmodavg)
#make mapview work on a mac
mapviewOptions(fgb = FALSE)
```

```{r load data}
#load data
records_2014 <- read.csv("./By_year/2014_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2015 <- read.csv("./By_year/2015_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2016 <- read.csv("./By_year/2016_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2017 <- read.csv("./By_year/2017_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2018 <- read.csv("./By_year/2018_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2019 <- read.csv("./By_year/2019_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2020 <- read.csv("./By_year/2020_Records_Yalbac_data.csv", stringsAsFactors = T)
```


```{r species names}
#figure out what each species name is 
unique(records_2014$Species)
unique(records_2015$Species)
unique(records_2016$species.name)
unique(records_2017$species.name)
unique(records_2018$species.name)
unique(records_2019$species.name)
unique(records_2020$species.name)
```

Ultimately need to make this:

unmarkedMultFrame(y, siteCovs, obsCovs, numPrimary, yearlySiteCovs)

y	- A matrix of the observed data.

siteCovs - Data frame of covariates that vary at the site level.

obsCovs	- Data frame of covariates that vary within site-year-observation level.

numPrimary - Number of primary time periods (seasons in the multiseason model).

yearlySiteCovs - Data frame containing covariates at the site-year level.

```{r 2014 cleaning}
#clean all 2014 data
records_2014_clean <- records_2014 %>%
  mutate(site = paste0(Cam.nr,Series))

unique(records_2014_clean$site)
```
71 cameras have observations for 2014. 79 cameras are listed with 3 not placed or stolen. Allegedly 76 should have been recording so logically 5 sites had 0 observations of any animal

These cameras that observed nothing (hypothetically) are 402A, 402C, 406F, 408F and 412C2.

length of all matrices for 2014 then should be 76 because there are than many sites.

```{r days op and ect}
#load operational days
cameraops2014 <- read.csv("./2014/OperationalDays2014.csv")

#data frame for 2014 
datafor2014 <- full_join(records_2014_clean, cameraops2014, by = "site")

#set as date column
datafor2014$startday <- as.Date(datafor2014$startday, "%Y-%m-%d")

#calculate end date
datafor2014_edit <- datafor2014 %>%
  mutate(endday = startday + daysop)

#find visit number
datafor2014_visit <- datafor2014_edit %>%
  mutate(obsdate = as.Date(paste0(Year,"-",Month,"-",Day), "%Y-%m-%d")) %>%
  mutate(visit = as.numeric(obsdate - startday) + 1)
```

