---
title: "Cleaning_data"
author: "Elise Boos"
date: '2022-08-15'
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load packages etc}
getwd()
#packages
require(tidyverse)
require(dplyr)
require(sf)
require(leaflet)
require(mapview)
require(readxl)
require(lubridate)
require(readxl)

library(unmarked)
library(AICcmodavg)
#make mapview work on a mac
mapviewOptions(fgb = FALSE)

#install flocker
#install.packages("remotes")
#remotes::install_github("jsocolar/flocker")

#remove.packages(c("StanHeaders", "rstan"))
#install.packages("StanHeaders", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
#install.packages("rstan", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))

#library(flocker)

#install.packages("camtrapR")
library(camtrapR)

```

```{r load data}
#load data
records_2014 <- read.csv("./By_year/2014_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2015 <- read.csv("./By_year/2015_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2016 <- read.csv("./By_year/2016_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2017 <- read.csv("./By_year/2017_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2018 <- read.csv("./By_year/2018_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2019 <- read.csv("./By_year/2019_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2020 <- read.csv("./By_year/2020_Records_Yalbac_data.csv", stringsAsFactors = T)
```

General notes:
some years cam name is 'site', later years cam name is 'Name'

```{r kelly species spelling cleaning}
#collect species names
names2014 <- unique(records_2014$Species)
names2015 <- unique(records_2015$Species)
names2016 <- unique(records_2016$species.name)
names2017 <- unique(records_2017$species.name)
names2018 <- unique(records_2018$species.name)
names2019 <- unique(records_2019$species.name)
names2020 <- unique(records_2020$species.name)


#save a string with ""
a <- paste0('"', paste(names2014, collapse='", "'), '"') #2014
print(a, quote=F)
b <- paste0('"', paste(names2015, collapse='", "'), '"') #2015
print(b, quote=F)
c <- paste0('"', paste(names2016, collapse='", "'), '"') #2016
print(c, quote=F)
d <- paste0('"', paste(names2017, collapse='", "'), '"') #2017
print(d, quote=F)
e <- paste0('"', paste(names2018, collapse='", "'), '"') #2018
print(e, quote=F)
f <- paste0('"', paste(names2019, collapse='", "'), '"') #2019
print(f, quote=F)
g <- paste0('"', paste(names2020, collapse='", "'), '"') #2020
print(g, quote=F)

##2014##########################################################################
################################################################################
#fix errors
records_2014$Species <- gsub("Squirel", "Squirrel", records_2014$Species)
records_2014$Species <- gsub("Black-faced Anthrush", "Black-faced Antthrush", records_2014$Species)
records_2014$Species <- gsub("Clay coloured Thrush", "Clay-coloured Thrush", records_2014$Species)
records_2014$Species <- gsub("Four-eyed Oppossum", "Four-eyed Opossum", records_2014$Species)
records_2014$Species <- gsub("Ruddy-quail Dove", "Ruddy Quail-dove", records_2014$Species)
records_2014$Species <- gsub("Stripe-nosed Skunk", "Striped Hog-nosed Skunk", records_2014$Species)

#fix misspelled species
records_2015$Species <- gsub("Collared Pecary", "Collared Peccary", records_2015$Species)
records_2015$Species <- gsub("Gray-necked woodrail ", "Gray-necked Woodrail", records_2015$Species)
records_2015$Species <- gsub("Great Black-Hawk", "Great Black Hawk", records_2015$Species)
records_2015$Species <- gsub("Louisiana waterthrush", "Louisiana Waterthrush", records_2015$Species)
records_2015$Species <- gsub("Ruddy-Quaildove", "Ruddy Quail-dove", records_2015$Species)
records_2015$Species <- gsub("Squirrel Cucko", "Squirrel Cuckoo", records_2015$Species)
records_2015$Species <- gsub("Squirrel Cuckooo", "Squirrel Cuckoo", records_2015$Species)
records_2015$Species <- gsub("Yucatan Squirel", "Yucatan Squirrel", records_2015$Species)
records_2015$Species <- gsub("Ornate Hawk-Eagle", "Ornate Hawk-eagle", records_2015$Species)

##2016##########################################################################
################################################################################
#fix misspelled species
records_2016$species.name <- gsub("Gray fox", "Gray Fox", records_2016$species.name)
records_2016$species.name <- gsub("Collared Pecary", "Collared Peccary", records_2016$species.name)
records_2016$species.name <- gsub("Crested guan", "Crested Guan", records_2016$species.name)
records_2016$species.name <- gsub("Clay coloured Thrush", "Clay-coloured Thrush", records_2016$species.name)
records_2016$species.name <- gsub("Commn Opossum", "Common Opossum", records_2016$species.name)
records_2016$species.name <- gsub("Gray hawk", "Gray Hawk", records_2016$species.name)
records_2016$species.name <- gsub("Rat and dragon fly", "Rat", records_2016$species.name)
records_2016$species.name <- gsub("Ruddy Quail Dove ", "Ruddy Quail-dove", records_2016$species.name)
records_2016$species.name <- gsub("Black hawk eagle", "Black Hawk-eagle", records_2016$species.name)
records_2016$species.name <- gsub("Roadside hawk", "Roadside Hawk", records_2016$species.name)
records_2016$species.name <- gsub("Eastern/Tropical pewee", "Pewee", records_2016$species.name)
records_2016$species.name <- gsub("agouti", "Agouti", records_2016$species.name)
records_2016$species.name <- gsub("agouti ", "Agouti", records_2016$species.name)
records_2016$species.name <- gsub("Agouti ", "Agouti", records_2016$species.name)
records_2016$species.name <- gsub("ocelot", "Ocelot", records_2016$species.name)

#fix errors
records_2017$species.name <- gsub("Gray fox", "Gray Fox", records_2017$species.name)
records_2017$species.name <- gsub("Ruddy Quail Dove ", "Ruddy Quail-dove", records_2017$species.name)
records_2017$species.name <- gsub("Collared Pecary", "Collared Peccary", records_2017$species.name)
records_2017$species.name <- gsub("Crested guan", "Crested Guan", records_2017$species.name)
records_2017$species.name <- gsub("Lesson's motmot", "Lesson's Motmot", records_2017$species.name)
records_2017$species.name <- gsub("Bare-throated Tiger-heron", "Bare-throated Tiger Heron", records_2017$species.name)
records_2017$species.name <- gsub("Black-crowned Night-Heron", "Black-crowned Night Heron", records_2017$species.name)
records_2017$species.name <- gsub("Great Blue-heron", "Great Blue Heron", records_2017$species.name)
records_2017$species.name <- gsub("Clay coloured Thrush", "Clay-coloured Thrush", records_2017$species.name)
records_2017$species.name <- gsub("Yellow-crowned Night-Heron", "Yellow-crowned Night Heron", records_2017$species.name)
records_2017$species.name <- gsub("Blue Ground-Dove", "Blue Ground Dove", records_2017$species.name)
records_2017$species.name <- gsub("Gray catbird", "Gray Catbird", records_2017$species.name)

##2018##########################################################################
################################################################################
#fix errors
records_2018$species.name <- gsub("Clay-colored Thrush", "Clay-coloured Thrush", records_2018$species.name)
records_2018$species.name <- gsub("Collared Pecary", "Collared Peccary", records_2018$species.name)
records_2018$species.name <- gsub("Crested guan", "Crested Guan", records_2018$species.name)
records_2018$species.name <- gsub("Gray fox", "Gray Fox", records_2018$species.name)
records_2018$species.name <- gsub("jaguar", "Jaguar", records_2018$species.name)
records_2018$species.name <- gsub("margay", "Margay", records_2018$species.name)
records_2018$species.name <- gsub("Ocellated turkey", "Ocellated Turkey", records_2018$species.name)
records_2018$species.name <- gsub("puma", "Puma", records_2018$species.name)
records_2018$species.name <- gsub("Ruddy Quail Dove ", "Ruddy Quail-dove", records_2018$species.name)
records_2018$species.name <- gsub("Striped hog-nosed skunk", "Striped Hog-nosed Skunk", records_2018$species.name)
records_2018$species.name <- gsub("tapir", "Tapir", records_2018$species.name)
records_2018$species.name <- gsub("tayra", "Tayra", records_2018$species.name)
records_2018$species.name <- gsub("Virginia opossum", "Virginia Opossum", records_2018$species.name)
records_2018$species.name <- gsub("Yellow-crowned Night-heron", "Yellow-crowned Night Heron", records_2018$species.name)
records_2018$species.name <- gsub("Bare-throated Tiger-heron", "Bare-throated Tiger Heron", records_2018$species.name)
records_2018$species.name <- gsub("Hooded warbler", "Hooded Warbler", records_2018$species.name)

##2019##########################################################################
################################################################################
#fix errors
records_2019$species.name <- gsub("Collared Pecary", "Collared Peccary", records_2019$species.name)
records_2019$species.name <- gsub("Gray fox", "Gray Fox", records_2019$species.name)
records_2019$species.name <- gsub("Virginia opossum", "Virginia Opossum", records_2019$species.name)
records_2019$species.name <- gsub("Crested guan", "Crested Guan", records_2019$species.name)
records_2019$species.name <- gsub("Ruddy Quail Dove ", "Ruddy Quail-dove", records_2019$species.name)
records_2019$species.name <- gsub("Clay coloured Thrush", "Clay-coloured Thrush", records_2019$species.name)
records_2019$species.name <- gsub("Blue Ground-Dove", "Blue Ground Dove", records_2019$species.name)
records_2019$species.name <- gsub("Red-throated ant-tanager", "Red-throated Ant-tanager", records_2019$species.name)
records_2019$species.name <- gsub("Bare-throated Tiger-heron", "Bare-throated Tiger Heron", records_2019$species.name)
records_2019$species.name <- gsub("White-tipped dove", "White-tipped Dove", records_2019$species.name)

##2020##########################################################################
################################################################################
#fix errors
records_2020$species.name <- gsub("Collared Pecary", "Collared Peccary", records_2020$species.name)
records_2020$species.name <- gsub("Blue Ground-Dove", "Blue Ground Dove", records_2020$species.name)
records_2020$species.name <- gsub("Gray fox", "Gray Fox", records_2020$species.name)
records_2020$species.name <- gsub("Ruddy Quail Dove ", "Ruddy Quail-dove", records_2020$species.name)
records_2020$species.name <- gsub("Clay coloured Thrush", "Clay-coloured Thrush", records_2020$species.name)
records_2020$species.name <- gsub("Crested guan", "Crested Guan", records_2020$species.name)

```

```{r mammal code - kelly}
records_2014$Mammal <- ifelse(records_2014$Species %in% c("Agouti", "Central American Tapir", "Collared Peccary", "Common Opossum", 
                                                          "Deppe's Squirrel", "Four-eyed Opossum", "Gray Fox", "Jaguar", "Jaguarundi", 
                                                          "Margay", "Mouse", "Nine-banded Armadillo", "Northern Tamandua", "Ocelot", "Paca", 
                                                          "Puma", "Racoon", "Rat", "Red Brocket Deer", "Rodent", "Squirrel", "Striped Hog-nosed Skunk", 
                                                          "Tayra", "White-lipped Peccary", "White-nosed Coati", "White-tailed Deer"), "Yes",
                       ifelse(records_2014$Species %in% c("0", "Ameiva undata", "Bare-throated Tiger Heron", "Black-faced Antthrush", "Blue-crowned Mot-mot", 
                                                 "Clay-coloured Thrush", "Collared Forest Falcon", "Common Pauraque", 
                                                 "Crested Guan", "White-tipped Dove", "Gray-headed Dove", "Gray-necked Woodrail", "Great Black Hawk", 
                                                 "Great Curassow", "Great Tinamou", "INDENT Dove", "INDENT Mammal", "INDENT Owl", "INDENT Pigeon", 
                                                 "Ocellated Turkey", "Ovenbird", "Plain Chachalaca", "Red-billed Pigeon", "Ruddy Quail-dove", 
                                                 "Slaty-breasted Tinamou", "Swainson's Trush", "Tinamou", "UNID", "White Hawk", "White-collard Seedeater", 
                                                 "Wood Thrush"), "No", NA))

#add mammal Y/N
records_2015$Mammal <- ifelse(records_2015$Species %in% c("Agouti", "Collared Peccary", "Deppe's Squirrel", "Gray Fox", 
                                                          "Jaguar", "Margay", "Mouse", "Nine-banded Armadillo", 
                                                          "Northern Tamandua", "Ocelot", "Opossum", "Paca", 
                                                          "Puma", "Racoon", "Red Brocket Deer", "Striped Hog-nosed Skunk",
                                                          "Stripe-throated Hermit", "Swainson's Thrush", "Tapir", "Tayra", 
                                                          "White-lipped Peccary", "White-nosed Coati", "White-tailed Deer", 
                                                          "Yucatan Squirrel"), "Yes",
                              ifelse(records_2015$Species %in% c("Bare-throated Tiger Heron", "Birds", "Black-faced Ant Thrush", 
                                                                                        "Clay coloured Thrush", "Blue Ground Dove", "Crested Guan", 
                                                                                        "Gray-headed Dove", "Gray-headed Kite", "Gray-necked Woodrail",
                                                                                        "Great Black Hawk", "Great Curassow", "Great Tinamou", "Green-backed Sparrow", 
                                                                                        "Hawk", "Louisiana Waterthrush", "Northern Waterthrush", "Ocellated Turkey", 
                                                                                        "Ornate Hawk-eagle", "Pauraque", "Plain Chachalaca", "Ruddy Quail-dove",  
                                                                                        "Slaty-breasted Tinamou", "Squirrel Cuckoo", "Thicket Tinamou", "UNID Tinamou", 
                                                                                        "White-tipped Dove", "Wood Thrush", ""), "No", NA))

#add mammal Y/N
records_2016$Mammal <- ifelse(records_2016$species.name %in% c("Agouti", "Gray Fox", "Margay", "Ocelot", "Puma", "Spider Monkey", 
                                                          "Tayra", "White-nosed Coati", "White-tailed Deer", "Collared Peccary", 
                                                          "Paca", "Red Brocket Deer", "Jaguar", "Common Opossum", "Rat", 
                                                          "Baird's Tapir", "White-lipped Peccary", "Striped Hog-nosed Skunk", "Racoon", 
                                                          "Brown Four-eyed Opossum", "Nine-banded Armadillo", "Tamandua", 
                                                          "Ocelot", "Rodent"), "Yes",
                              ifelse(records_2016$species.name %in% c("Great Curassow", "unknown", "Ocellated Turkey", "Crested Guan", 
                                                                 "Chachalaka", "Clay-coloured Thrush", "Dragon fly", "Flycatcher", 
                                                                 "Gray Hawk", "Gray-headed Dove", "Great Tinamou", "lizard", 
                                                                 "Ruddy Quail-dove", "Scaled Pigeon", "White-tipped Dove", 
                                                                 "Slaty-breasted Tinamou", "Wood Thrush", "Great Black Hawk", 
                                                                 "Limpkin", "Northern Waterthrush", "Black Hawk-eagle",
                                                                 "Blue Ground Dove", "Roadside Hawk", "Pewee", "Boat-billed Heron", 
                                                                 "Little Tinamou", ""), "No", NA))

#add mammal Y/N
records_2018$Mammal <- ifelse(records_2018$species.name %in% c("Agouti", "Baird's Tapir", "Collared Peccary", "Common Opossum", 
                                                               "Gray Four-eyed Opossum", "Gray Fox", "Jaguar", "Jaguarundi", 
                                                               "Margay", "Nine-banded Armadillo", "Northern Raccoon", "Ocelot", "Paca", 
                                                               "Puma", "Red Brocket Deer", "Striped Hog-nosed Skunk", "Tamandua", 
                                                               "Tapir", "Tayra", "Unidentified Mammal", "Unknown Rodent", "Virginia Opossum",
                                                               "White-nosed Coati", "White-tailed Deer", "Yucatan Squirrel", "Unidentified mammal", 
                                                               "White-lipped Peccary"), "Yes",
                              ifelse(records_2018$species.name %in% c("Chachalaca", "Clay-coloured Thrush", "Crested Guan", "Gray-headed Dove", 
                                                                      "Great Black Hawk", "Great Curassow", "Great Tinamou", "Lineated Woodpecker", 
                                                                      "Mottled Owl", "Ocellated Turkey", "Pauraque", "Ruddy Quail-dove", 
                                                                      "Russet-naped Woodrail", "Slaty-breasted Tinamou", "Tinamou sp.", "Unidentified ", 
                                                                      "Wood Thrush", "Yellow-crowned Night Heron", "Bare-throated Tiger Heron", 
                                                                      "Hooded Warbler", "Nothern Waterthrush", "Red-throated Ant-tanager", 
                                                                      "Ruddy Ground-dove", "Summer Tanager", "Thicket Tinamou", "Unknown", "Unknown bird", 
                                                                      ""), "No", NA))

records_2019$Mammal <- ifelse(records_2019$species.name %in% c("Agouti", "White-tailed Deer", "Deppe's Squirrel", "Collared Peccary", 
                                                               "Puma", "Striped Hog-nosed Skunk", "White-lipped Peccary", "Ocelot", 
                                                               "Gray Fox", "Common Opossum", "Virginia Opossum", "White-nosed Coati", 
                                                               "Brown Four-eyed Opossum", "Red Brocket Deer", "Baird's Tapir", 
                                                               "Nine-banded Armadillo", "Jaguar", "Tayra", "Paca", "Racoon", "Margay", 
                                                               "Northern Tamandua", "Bat", "Rodent", "Jaguarundi", "Coyote", "Unknown mammal"), "Yes",
                              ifelse(records_2019$species.name %in% c("Great Curassow", "Ocellated Turkey", "Wood Thrush", "Gray-headed Dove", 
                                                                      "Black-faced Antthrush", "Great Tinamou", "Crested Guan", "Ruddy Quail-dove", 
                                                                      "Clay-coloured Thrush", "Slaty-breasted Tinamou", "Blue Ground Dove", 
                                                                      "Plain Chachalaca", "Unknown bird", "Red-throated Ant-tanager", "Great Black Hawk", 
                                                                      "Russet-naped Woodrail", "Bare-throated Tiger Heron", "Boat-billed Heron", 
                                                                      "Swainson's Thrush", "Common Pauraque", "Northern Waterthrush", "White-tipped Dove", 
                                                                      "Roadside Hawk", "Ornate Hawk-eagle", "Owl", "Blue Grosbeak",  "Green Heron"), "No", NA))

################################################################################

#add mammal Y/N
records_2020$Mammal <- ifelse(records_2020$species.name %in% c("White-lipped Peccary", "White-nosed Coati", "Collared Peccary", "Baird's Tapir", 
                                                               "Paca", "Red Brocket Deer", "White-tailed Deer", "Jaguar", "Gray Fox", 
                                                               "Racoon", "Agouti", "Puma", "Squirrel", "Rodent", "Mammal", "Virginia Opossum", 
                                                               "Ocelot", "Tayra", "Common Opossum", "Margay", "Northern Tamandua", "Striped Hog-nosed Skunk", 
                                                               "Nine-banded Armadillo"), "Yes",
                              ifelse(records_2020$species.name %in% c("Great Curassow", "Ocellated Turkey", "Blue Ground Dove", "Gray-headed Dove", 
                                                                      "Great Tinamou", "Slaty-breasted Tinamou", "Little Tinamou", "Plain Chachalaca", 
                                                                      "Ruddy Quail-dove", "Swainson's Thrush", "Toad", "Clay-coloured Thrush", "Green-backed Sparrow", 
                                                                      "Hooded Warbler", "Brown-headed Flycatcher", "Ameiva", "Crested Guan", 
                                                                      "Common Pauraque", "Wood Thrush", "Bird", ""), "No", NA))


```

Ultimately need to make this:

unmarkedMultFrame(y, siteCovs, obsCovs, numPrimary, yearlySiteCovs)

y	- A matrix of the observed data.

siteCovs - Data frame of covariates that vary at the site level.

obsCovs	- Data frame of covariates that vary within site-year-observation level.

numPrimary - Number of primary time periods (seasons in the multiseason model).

yearlySiteCovs - Data frame containing covariates at the site-year level.

```{r 2014 cleaning}
#clean all 2014 data
records_2014_clean <- records_2014 %>%
  mutate(site = paste0(Cam.nr,Series))

#unique(records_2014_clean$site)
```
71 cameras have observations for 2014. 79 cameras are listed with 3 not placed or stolen. Allegedly 76 should have been recording so logically 5 sites had 0 observations of any animal

These cameras that observed nothing or didn't work (hypothetically) are 402A, 402C, 406F, 408F and 412C2.

length of all matrices for 2014 then should be 76 because there are than many sites.

```{r days op and ect}
#load operational days
#LOAD IN OTHER VARIABLES FOR 2014 LIKE LOGGING ETC
cameraops2014 <- read.csv("./2014/Cameras_2014.csv", header = T)

#data frame for 2014 
datafor2014 <- full_join(records_2014_clean, cameraops2014, by = c("site"="Name"))

#set as date column
datafor2014$Date.Placement <- as.Date(datafor2014$Date.Placement, "%B %d, %Y")
datafor2014$Date.Removed <- as.Date(datafor2014$Date.Removed, "%Y-%m-%d")

#calculate end date
datafor2014_edit <- datafor2014 %>%
  mutate(Last.Record = Date.Placement + Days +1)
#add one here?

#fix error for 401A picture 28 
datafor2014_edit[475, ]$Month <- 3
datafor2014_edit[475, ]$Day <- 24

#find visit number
datafor2014_visit <- datafor2014_edit %>%
  mutate(obsdate = as.Date(paste0(Year,"-",Month,"-",Day), "%Y-%m-%d")) %>%
  mutate(visit = as.numeric(obsdate - Date.Placement +1))

#error finder 
datafor2014_errors <- datafor2014_visit %>%
  mutate(errors = ifelse(obsdate > Last.Record, TRUE, FALSE)) %>%
  mutate(opp_errors = ifelse(obsdate < Date.Placement, TRUE, FALSE))
#error summary
summary(datafor2014_errors$errors)
summary(datafor2014_errors$opp_errors)

#last record update for after typos fixed
fixerrors_2014 <- datafor2014_errors %>%
  mutate(max.date = if_else(as.Date(obsdate) > as.Date(Last.Record), obsdate, Last.Record)) %>%
  group_by(site) %>%
  summarize(Last.Record_clean = max(max.date))

datafor2014_final <- full_join(datafor2014_errors, fixerrors_2014, by = "site")

#check final
datafor2014_check <- datafor2014_final %>%
   mutate(errors_final = ifelse(obsdate > Last.Record_clean, TRUE, FALSE))

#clean up for final dataframe
data2014_final <- datafor2014_final %>%
  select(site, 1:15,27, obsdate, Last.Record_clean,Date.Placement, Date.Removed, Functioning., visit, Mammal,Days)

data2014_final <- data2014_final  %>%
  mutate(date_time_obs = paste(obsdate, sprintf("%.2f", round(Hour_exact, 2)))) %>%
  drop_na(obsdate) 

camdataprep_2014 <- data2014_final %>%
  select(site, Date.Placement, Days, Last.Record_clean, Northing.x, Easting.x)
camopdataframe2014 <- camdataprep_2014[!duplicated(camdataprep_2014$site), ]

write.csv(data2014_final, "data2014_final.csv", row.names = FALSE)
write.csv(camopdataframe2014, "camop_2014.csv", row.names = FALSE)

covs2014 <- read.csv("./2014/Covariates_Cams_2014.csv", stringsAsFactors = TRUE)
sitecovs2014 <- left_join(camopdataframe2014, cameraops2014, c("site"="Name"))
sitecovs2014 <- left_join(sitecovs2014, covs2014, c("site"="Name"))

write.csv(sitecovs2014, "sitecovs2014.csv", row.names = FALSE)
```
Errors in data found that need to fix: 

410C2 is repeated 410C needed to be fixed in the days operational in the original excel file.

Record 2234 is an obs after the camera was removed so need to removed probably. A couple data points are like this need to look into!!!! 241 observations are after end date!
ERROR SEEMS TO BE IN C2 and C points - repeated in data of days operational (more than this tho)

are days operational including day 1? From looking at data day 1 is included as one operational day. Actually idk

Some points that are inconsistent going to need to just put 0s to the days the camera was out there.

39 or more like 47? cameras have incorrect operational days. for 2014 might just have to do the days the camera was out

~~~~~~~~~
New notes 

record 475 for 401A i believe the month and day are flipped 

still need to figure out where the error is in days cam out 
------------------------------
```{r 2015 cleaning}
#clean all 2014 data
records_2015_clean <- records_2015 %>%
  mutate(site = paste0(Cam.nr,Series)) %>%
  drop_na(Cam.nr)
records_2015_clean[records_2015_clean$site == "511G",]$site <- "511FG"

#unique(records_2015_clean$site)
```

```{r joining cams}
cameraops2015 <- read.csv("./2015/Cameras_2015.csv", header = T)

datafor2015 <- full_join(records_2015_clean, cameraops2015, by = c("site" = "Name"))

#set as date 
datafor2015$Date_Place <- as.Date(datafor2015$Date_Place, "%B %d, %Y")
datafor2015$Date_Remov <- as.Date(datafor2015$Date_Remov, "%B %d, %Y")

#picture 251 510E 9 for month instead of 10
#wrong placement day for 508E
datafor2015[2043, ]$Month <- 9
datafor2015[datafor2015$site == "508E",]$Date_Place <- as.Date("7/29/2015", "%m/%d/%Y") 
datafor2015[2066, ]$Month <- 12
datafor2015[2066, ]$Day <- 11

datafor2015_edit <- datafor2015 %>%
  mutate(Last.Record = Date_Place + Days +1)
#add one here?

#find visit number
datafor2015_visit <- datafor2015_edit %>%
  mutate(obsdate = as.Date(paste0(Year,"-",Month,"-",Day), "%Y-%m-%d")) %>%
  mutate(visit = as.numeric((obsdate - Date_Place) + 1))

#error finder 
datafor2015_errors <- datafor2015_visit %>%
  mutate(errors = ifelse(obsdate > Last.Record, TRUE, FALSE)) 
summary(datafor2015_errors$errors)

#507H has typo 2016 instead of 2015 
datafor2015_errorsfixed <- datafor2015_errors %>%
  mutate(year_clean = ifelse(site == "507H" & errors == TRUE, paste(2015), Year))


#redo visit and errors 
datafor2015_visit <- datafor2015_errorsfixed %>%
  mutate(obsdate = as.Date(paste0(year_clean,"-",Month,"-",Day), "%Y-%m-%d")) %>%
  mutate(visit = as.numeric((obsdate - Date_Place) + 1))

datafor2015_errors <- datafor2015_visit %>%
  mutate(errors = ifelse(obsdate > Last.Record, TRUE, FALSE)) %>%
  mutate(opp_errors = ifelse(obsdate < Date_Place, TRUE, FALSE)) 
summary(datafor2015_errors$errors)
summary(datafor2015_errors$opp_errors)

#clean up for final dataframe
data2015_final <- datafor2015_visit %>%
  select(site, 1:3,site,6:15,17,18,20,22, Date_Place,Date_Remov,obsdate, Last.Record, Days, visit, Mammal, 31,33:39) %>%
  mutate(date_time_obs = paste(obsdate, sprintf("%.2f", round(Hour_exact, 2))))

nonfunc2015 <- data2015_final[data2015_final$Days == 0,]
rows <- rownames(nonfunc2015)
data2015_final <- data2015_final[-as.numeric(rows),]

camdataprep_2015 <- data2015_final %>%
  select(site, Date_Place, Days, Last.Record, Northing.1, Easting.1)
camopdataframe2015 <- camdataprep_2015[!duplicated(camdataprep_2015$site), ]

data2015_final <- data2015_final %>%
  drop_na(obsdate)

#make the 3 dataframes 

write.csv(camopdataframe2015, "camop_2015.csv", row.names = FALSE)
write.csv(data2015_final, "data2015_final.csv", row.names = FALSE)

#camop2015 <- cameraOperation(camopdataframe2015, stationCol = "site", setupCol = "Date_Place", retrievalCol = "Last.Record", occasionStartTime = 0 , dateFormat = "%Y-%m-%d", writecsv = FALSE)
sitecovs2015 <- left_join(camopdataframe2015, cameraops2015, c("site"="Name"))
write.csv(sitecovs2015, "sitecovs2015.csv", row.names = FALSE)
```

Only errors in 7 of the observations - can check back for typos 

--------------------
```{r cleaning 2016}
records_2016_clean <- records_2016 %>%
  mutate(site = paste0(Cam.nr,series)) %>%
  drop_na(Cam.nr)

records_2016_clean[records_2016_clean$site %in% c("607Ab","607Ac", "607Ad"),]$site <- "607A"

unique(records_2016_clean$site)

```

```{r joining cam data}
cameraops2016 <- read.csv("./2016/Cameras_2016.csv", header = T)

#issue with colname import for kelly
#colnames(cameraops2016)[colnames(cameraops2016)=="ï..Name"] <- "Name"

unique(cameraops2016$Name)


datafor2016 <- full_join(records_2016_clean, cameraops2016, by = c("site" = "Name"))

#set as date 
datafor2016$Date.Placement <- as.Date(datafor2016$Date.Placement, "%B %d, %Y")
datafor2016$Date.Removed <- as.Date(datafor2016$Date.Removed, "%B %d, %Y")



#calc end date for cam
datafor2016_edit <- datafor2016 %>%
  mutate(Last.Record = Date.Placement + Days + 1)

datafor2016_edit[1546,]$Last.Record <- "2016-04-12"
#remove 3 records from 2012 
#remove 4078 for 602J picture doesn't exist

datafor2016_edit[1225, ]$Month <- 6
datafor2016_edit[1232, ]$Month <- 6
datafor2016_edit[1225, ]$Day <- 10
datafor2016_edit[19, ]$Hour.Exact <- 9.36

#604 K <- day and night for hour!!!!!
no_hour <- datafor2016_edit[datafor2016_edit$Comment == "Date not working",]
no_hour <- no_hour %>%
  drop_na(site)
rows_no_hour <- rownames(no_hour)
ronumb <- as.numeric(rows_no_hour)
datafor2016_new <- datafor2016_edit[-c(1312,1313,1319,102,928,ronumb), ]
#datafor2016_edit <- datafor2016_edit[-c(434:458),]


#visit day
datafor2016_visit <- datafor2016_new %>%
  mutate(obsdate = as.Date(paste0(Year,"-",Month,"-",Day), "%Y-%m-%d")) %>%
  mutate(visit = as.numeric((obsdate - Date.Placement) + 1))

#error finder 
datafor2016_errors <- datafor2016_visit %>%
  mutate(errors = ifelse(obsdate > Last.Record, TRUE, FALSE)) %>%
  mutate(opp_errors = ifelse(obsdate < Date.Placement, TRUE, FALSE)) 
summary(datafor2016_errors$errors)
summary(datafor2016_errors$opp_errors)


data2016_final <- datafor2016_errors  %>%
  mutate(date_time_obs = paste(obsdate, sprintf("%.2f",round(as.numeric(as.character(Hour.Exact)),2)))) %>%
  select(4:11,13:21,25:35,38)

nonfunc2016 <- data2016_final[data2016_final$Days == 0,]
sites <- nonfunc2016$site
data2016_final <- data2016_final[!(data2016_final$site %in% sites),]

camdataprep_2016 <- data2016_final %>%
  select(site, Date.Placement, Days, Last.Record, Northing, Easting)
camopdataframe2016 <- camdataprep_2016[!duplicated(camdataprep_2016$site), ]

data2016_final <- data2016_final%>%
  drop_na(obsdate)


write.csv(camopdataframe2016, "camop_2016.csv", row.names = FALSE)
write.csv(data2016_final, "data2016_final.csv", row.names = FALSE)

#years incorrect some marked as 2015 instead of 2016
#fixed in excel

sitecovs2016 <- left_join(camopdataframe2016, cameraops2016, c("site"="Name"))

write.csv(sitecovs2016, "sitecovs2016.csv", row.names = FALSE)
```

607A -> subsetted into Ab Ac Ad, not sure if this is just the location of 607A or what the reason for subsetting is. Accounts for all the NA's.

Some january placement days were listed as 2015 instead of 2016 -> once that is fixed only 5 errors

284 NA's = the 607A deviations

Maybe check photo to see observation day. 

----------------------

```{r cleaning 2017}
records_2017_clean <- records_2017 %>%
  mutate(site = paste0(Cam.nr,series)) %>%
  drop_na(Cam.nr)

unique(records_2017_clean$site)

```

```{r matching 2017 camera data}
cameraops2017 <- read.csv("./2017/Cameras_2017.csv", header = T)

#issue with colname import
#colnames(cameraops2017)[colnames(cameraops2017)=="ï..Name"] <- "Name"

unique(cameraops2017$Name)

datafor2017 <- full_join(records_2017_clean, cameraops2017, by = c("site" = "Name"))

#set as date 
datafor2017$Date.Placement <- as.Date(datafor2017$Date.Placement, "%d-%b-%y")
datafor2017$Date.Removed <- as.Date(datafor2017$Date.Removed, "%d-%b-%y")

#calc end date for cam
datafor2017_edit <- datafor2017 %>%
  mutate(Last.Record = Date.Placement + Days + 1)

#fix obs date for 10260313 in 710E
datafor2017_edit[1248, ]$Day <- 26

#visit day
datafor2017_visit <- datafor2017_edit %>%
  mutate(obsdate = as.Date(paste0(Year,"-",Month,"-",Day), "%Y-%m-%d")) %>%
  mutate(visit = as.numeric((obsdate - Date.Placement) + 1))

#error finder 
datafor2017_errors <- datafor2017_visit %>%
  mutate(errors = ifelse(obsdate > Last.Record, TRUE, FALSE)) %>%
  mutate(opp_errors = ifelse(obsdate < Date.Placement, TRUE, FALSE)) 
summary(datafor2017_errors$errors)
summary(datafor2017_errors$opp_errors)

#fix errors 

#last record update for after typos fixed
fixerrors_2017 <- datafor2017_errors %>%
  mutate(max.date = if_else(as.Date(obsdate) > as.Date(Last.Record), obsdate, Last.Record)) %>%
  group_by(site) %>%
  summarize(Last.Record_clean = max(max.date)) 
  
datafor2017_int <- full_join(datafor2017_errors, fixerrors_2017, by = "site")

fixerrors_2017_cont <- datafor2017_int %>%
  mutate(min.date = if_else(as.Date(obsdate) < as.Date(Date.Placement), obsdate, Date.Placement)) %>%
  group_by(site) %>%
  summarize(First_Obs = min(min.date))

datafor2017_final <- full_join(datafor2017_int, fixerrors_2017_cont, by = "site")

#check final
datafor2017_check <- datafor2017_final %>%
  mutate(errors_final = ifelse(obsdate > Last.Record_clean, TRUE, FALSE)) %>%
  mutate(errors_final2 = ifelse(obsdate < First_Obs, TRUE, FALSE))

data2017_final <- datafor2017_final  %>%
  mutate(visit = as.numeric((obsdate - First_Obs + 1))) %>%
  select(site, 1:14, 17,18,34,19:21,25:33, First_Obs, Last.Record_clean) %>%
  mutate(date_time_obs = paste(obsdate, sprintf("%.2f", round(Hour.Exact, 2))))
#in this case all had 0 days 
data2017_final <- data2017_final %>%
  drop_na(Hour.Exact)

camdataprep_2017 <- data2017_final %>%
  select(site, First_Obs, Days, Last.Record_clean, Northing, Easting, Logging)
camopdataframe2017 <- camdataprep_2017[!duplicated(camdataprep_2017$site), ]




write.csv(camopdataframe2017, "camop_2017.csv", row.names = FALSE)
write.csv(data2017_final, "data2017_final.csv", row.names = FALSE)

sitecovs2017 <- left_join(camopdataframe2017, cameraops2017, c("site"="Name"))

write.csv(sitecovs2017, "sitecovs2017.csv", row.names = FALSE)
```

Errors with cam 710E (maybe other cams too) Have to do with "Last.Record". This date is different than the date removed. Date removed would not result in these errors.

Edit clean data frame to have first_obs and last.record_clean to be date.placement and last.record

```{r making occupancy dataframe 2017}

camop2017 <- cameraOperation(camopdataframe2017, stationCol = "site", setupCol = "First_Obs", retrievalCol = "Last.Record_clean", occasionStartTime = 0 , dateFormat = "%Y-%m-%d", writecsv = FALSE)


detect_hist_2017_grayfox <- detectionHistory(recordTable = data2017_final, species = "Gray Fox", camOp= camop2017, output = "binary", stationCol = "site", speciesCol = "species.name", day1 ="station", recordDateTimeCol = "date_time_obs", recordDateTimeFormat =  "%Y-%m-%d %H.%M", timeZone = "UTC", occasionLength = 1, includeEffort = TRUE, scaleEffort = FALSE, writecsv = FALSE)

detectionhist <- detect_hist_2017_grayfox$detection_history
```
In making the prep data frames need to make sure using correct date placements ie the calculated ones. 

```{r model test}
covariates <- read_xlsx("Covariates_cams.xlsx")

cams700 <- covariates[188:263,]

dataframe_det <- as.data.frame(detectionhist)
dataframe_det$rn <- row.names(dataframe_det)


join <- left_join(dataframe_det,cams700, by = c("rn" = "CameraID"))
cams700_clean <- select(join, 116:215)

unmarkedFrame <- unmarkedFrameOccu(detectionhist, siteCovs = cams700_clean)
summary(unmarkedFrame)

#flock_data <- make_flocker_data(obs = detectionhist, unit_covs = cams700_clean)
```

```{r ooo model}
#flock(f_occ = ~ 1, f_det = ~ 1, flocker_data = flock_data)

fm <- occu(formula = ~1
                     ~1,
           data = unmarkedFrame)
summary(fm)

backTransform(fm, type = "state")
backTransform(fm, type = "det")

#try a covariate

fm2 <- occu(formula = ~1
                     ~ Avg_pr_2017,
           data = unmarkedFrame)
summary(fm2)
```

----------------------

```{r cleaning 2018}
records_2018_clean <- records_2018 %>%
  mutate(series_caps = toupper(series)) %>%
  mutate(site = paste0(Cam.nr,series_caps)) %>%
  drop_na(Cam.nr)

records_2018_clean$Year <- 2018

unique(records_2018_clean$site)
```

```{r matching 2018 cam data}
cameraops2018 <- read.csv("./2018/Cameras_2018.csv", header = T)

unique(cameraops2018$Name)

datafor2018 <- full_join(records_2018_clean, cameraops2018, by = c("site" = "Name"))

#set as date 
datafor2018$Date.Placement <- as.Date(datafor2018$Date.Placement, "%d-%b-%y")
datafor2018$Date.Removed <- as.Date(datafor2018$Date.Removed, "%d-%b-%y")
datafor2018$Last.record <- as.Date(datafor2018$Last.record, "%d-%b-%y")

#visit day
datafor2018_visit <- datafor2018 %>%
  mutate(obsdate = as.Date(paste0(Year,"-",Month,"-",Day), "%Y-%m-%d")) %>%
  mutate(visit = as.numeric((obsdate - Date.Placement) + 1))

#error finder 
datafor2018_errors <- datafor2018_visit %>%
  mutate(errors = ifelse(obsdate > Last.record, TRUE, FALSE)) %>%
  mutate(opp_errors = ifelse(obsdate < Date.Placement, TRUE, FALSE)) 
summary(datafor2018_errors$errors)


#fix errors 
#last record update for after typos fixed
fixerrors_2018 <- datafor2018_errors %>%
  mutate(max.date = if_else(as.Date(obsdate) > as.Date(Last.record), obsdate, Last.record)) %>%
  group_by(site) %>%
  summarize(Last.Record_clean = max(max.date)) 
  
datafor2018_int <- full_join(datafor2018_errors, fixerrors_2018, by = "site")

fixerrors_2018_cont <- datafor2018_int %>%
  mutate(min.date = if_else(as.Date(obsdate) < as.Date(Date.Placement), obsdate, Date.Placement)) %>%
  group_by(site) %>%
  summarize(First_Obs = min(min.date))

datafor2018_final <- full_join(datafor2018_int, fixerrors_2018_cont, by = "site")

#check final
datafor2018_check <- datafor2018_final %>%
  mutate(errors_final = ifelse(obsdate > Last.Record_clean, TRUE, FALSE)) %>%
  mutate(errors_final2 = ifelse(obsdate < First_Obs, TRUE, FALSE))
summary(datafor2018_check$errors_final)
summary(datafor2018_check$errors_final2)

data2018_final <- datafor2018_final  %>%
  mutate(visit = as.numeric((obsdate - First_Obs + 1))) %>%
  select(site,2,7:15,17:18,21,23,Easting.y,Northing.y,34,35, obsdate, First_Obs, Last.Record_clean, Days) %>%
  mutate(date_time_obs = paste(obsdate, sprintf("%.2f", round(Hour.Exact, 2))))
#in this case all had 0 days 
data2018_final <- data2018_final %>%
  drop_na(Hour.Exact)

camdataprep_2018 <- data2018_final %>%
  select(site, First_Obs, Days, Last.Record_clean, Northing.y, Easting.y)
camopdataframe2018 <- camdataprep_2018[!duplicated(camdataprep_2018$site), ]



write.csv(camopdataframe2018, "camop_2018.csv", row.names = FALSE)
write.csv(data2018_final, "data2018_final.csv", row.names = FALSE)

sitecovs2018 <- left_join(camopdataframe2018, cameraops2018, c("site"="Name"))
write.csv(sitecovs2018, "sitecovs2018.csv", row.names = FALSE)
```
NOTE: remove non functions cameras maybe and also remove the days record

CHECK BACK AT PHOTOS to see if obs date is right!
--------------------

```{r cleaning 2019}
unique(records_2019$Camera)

records_2019_clean <- records_2019 %>%
  mutate(site = paste0(Cam.nr,series)) %>%
  drop_na(Cam.nr)

```


```{r matching 2019 cam data}
cameraops2019 <- read.csv("./2019/Cameras_2019.csv", header = T)

unique(cameraops2019$Name)

datafor2019 <- full_join(records_2019_clean, cameraops2019, by = c("site" = "Name"))

#set as date 
datafor2019$Date.Placement <- as.Date(as.character(datafor2019$Date.Placement), "%Y%m%d")
datafor2019$Date.Removed <- as.Date(as.character(datafor2019$Date.Removed), "%Y%m%d")
datafor2019$Last.record <- as.Date(as.character(datafor2019$Last.record), "%Y%m%d")

datafor2019[datafor2019$site == "907E",]$Last.record <- as.Date("2019-08-24","%Y-%m-%d")
#visit day
datafor2019_visit <- datafor2019 %>%
  mutate(obsdate = as.Date(paste0(Year,"-",Month,"-",Day), "%Y-%m-%d")) %>%
  mutate(visit = as.numeric((obsdate - Date.Placement) + 1))

#error finder 
datafor2019_errors <- datafor2019_visit %>%
  mutate(errors = ifelse(obsdate > Last.record, TRUE, FALSE)) %>%
  mutate(opp_errors = ifelse(obsdate < Date.Placement, TRUE, FALSE))
summary(datafor2019_errors$errors)
summary(datafor2019_errors$opp_errors)

```
909A month is wrong 

series wrong for 905E (says F) - something wrong here, not sure if location is right maybe is F 

only two cameras with issues XXXX

remove empty rows
----------------

```{r cleaning 2020}

unique(records_2020$Name)

```

```{r matching 2020 cam data}
cameraops2020 <- read.csv("./2020/Cameras_2020.csv", header = T)

unique(cameraops2020$Name)

datafor2020 <- full_join(records_2020, cameraops2020, by = "Name")

#set as date 
datafor2020$Date.Placement <- as.Date(as.character(datafor2020$Date.Placement), "%Y%m%d")
datafor2020$Date.Removed <- as.Date(as.character(datafor2020$Date.Removed), "%Y%m%d")
datafor2020$Last.record <- as.Date(as.character(datafor2020$Last.record), "%Y%m%d")

#visit day
datafor2020_visit <- datafor2020 %>%
  mutate(obsdate = as.Date(paste0(Year,"-",Month,"-",Day), "%Y-%m-%d")) %>%
  mutate(visit = as.numeric((obsdate - Date.Placement) + 1))

#error finder 
datafor2020_errors <- datafor2020_visit %>%
  mutate(errors = ifelse(obsdate > Last.record, TRUE, FALSE)) %>%
  mutate(opp_errors = ifelse(obsdate < Date.Placement, TRUE, FALSE)) 
summary(datafor2020_errors$errors)
summary(datafor2020_errors$opp_errors)

```
weird stuff with 2020 some obs dates are before the camera was placed.


Create a spatial data frame for the camaeras for all the years:
```{r combining cameras and locationas}
cameraops2014_edit <- cameraops2014 %>%
  select(Name,Easting,Northing) %>%
  mutate(Record_Year = 2014)
cameraops2015_edit<-cameraops2015 %>%
  select(Name,Easting,Northing) %>%
  mutate(Record_Year = 2015) 
cameraops2016_edit<-cameraops2016 %>%
  select(Name,Easting,Northing) %>%
  mutate(Record_Year = 2016) 
cameraops2017_edit<-cameraops2017 %>%
  select(Name,Easting,Northing) %>%
  mutate(Record_Year = 2017)  
cameraops2018_edit<-cameraops2018 %>%
  select(Name,Easting,Northing) %>%
  mutate(Record_Year = 2018) 
cameraops2019_edit<-cameraops2019 %>%
  select(Name,Easting,Northing) %>%
  mutate(Record_Year = 2019) 
cameraops2020_edit<-cameraops2020 %>%
  select(Name,Easting,Northing) %>%
  mutate(Record_Year = 2020) 
cameras_all<- rbind(cameraops2014_edit,cameraops2015_edit,cameraops2016_edit,cameraops2017_edit,cameraops2018_edit,cameraops2019_edit,cameraops2020_edit)
cameras_all_clean <- drop_na(cameras_all)

```

```{r mapview}
cameras_all.sf <- cameras_all_clean %>%
  st_as_sf(coords = c('Easting', 'Northing'),
         crs = 26716)
mapview(cameras_all.sf, zcol = "Record_Year")



st_write(cameras_all.sf, "cams2014to2020_withyear.shp", driver = "ESRI Shapefile")
```

