---
title: "GrayFoxOccupancy"
author: "Elise Boos"
date: '2022-08-05'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#set wd 
getwd()
```

## Gray fox occupancy

```{r load packages etc}
getwd()
#packages
require(tidyverse)
require(dplyr)
require(sf)
require(leaflet)
require(mapview)
require(readxl)

library(unmarked)
library(AICcmodavg)
#make mapview work on a mac
mapviewOptions(fgb = FALSE)
```

```{r load data}
#load data
records_2014 <- read.csv("./By_year/2014_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2015 <- read.csv("./By_year/2015_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2016 <- read.csv("./By_year/2016_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2017 <- read.csv("./By_year/2017_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2018 <- read.csv("./By_year/2018_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2019 <- read.csv("./By_year/2019_Records_Yalbac_data.csv", stringsAsFactors = T)
records_2020 <- read.csv("./By_year/2020_Records_Yalbac_data.csv", stringsAsFactors = T)
```


```{r species names}
#figure out what each species name is 
unique(records_2014$Species)
unique(records_2015$Species)
unique(records_2016$species.name)
unique(records_2017$species.name)
unique(records_2018$species.name)
unique(records_2019$species.name)
unique(records_2020$species.name)
```

```{r filter data}
grayfox2014 <- records_2014 %>%
  filter(Species == "Gray Fox")

grayfox2015 <- records_2015 %>%
  filter(Species == "Gray Fox")

grayfox2016 <- records_2016 %>%
  filter(species.name == "Gray fox")

grayfox2017 <- records_2017 %>%
  filter(species.name == "Gray fox")

grayfox2018 <- records_2018 %>%
  filter(species.name == "Gray fox")

grayfox2019 <- records_2019 %>%
  filter(species.name == "Gray fox")

grayfox2020 <- records_2020 %>%
  filter(species.name == "Gray fox")
```

Ultimately need to make this:

unmarkedMultFrame(y, siteCovs, obsCovs, numPrimary, yearlySiteCovs)

y	- A matrix of the observed data.

siteCovs - Data frame of covariates that vary at the site level.

obsCovs	- Data frame of covariates that vary within site-year-observation level.

numPrimary - Number of primary time periods (seasons in the multiseason model).

yearlySiteCovs - Data frame containing covariates at the site-year level.

```{r 2014 cleaning}
#clean gray fox only data
grayfox2014_clean <- grayfox2014 %>%
  mutate(site = paste0(Cam.nr,Series))
  
unique(grayfox2014_clean$site)

#clean all 2014 data
records_2014_clean <- records_2014 %>%
  mutate(site = paste0(Cam.nr,Series))

unique(records_2014_clean$site)
```
71 cameras have observations for 2014. 79 cameras are listed with 3 not placed or stolen. Allegedly 76 should have been recording so logically 5 sites had 0 observations of any animal

These cameras that observed nothing (hypothetically) are 402A, 402C, 406F, 408F and 412C2.

length of all matrices for 2014 then should be 76 because there are than many sites.

```{r make observation data frame}
grayfoxobs2014 = data.frame(matrix(nrow = 71, ncol = 63))

recsites2014 <- unique(records_2014_clean$site)
#NoObs2014 <- c("402A", "402C", "406F", "408F", "412C2")
#sites2014 <- append(recsites2014, NoObs2014)

grayfoxobs2014[,1] <- recsites2014

#Camera days of operation for 2014
cameraops2014 <- read.csv("OperationsDays2014.csv")

#join the length of observation to dataframe
operationaldays <- full_join(grayfoxobs2014, cameraops2014, by = c("X1" = "site"))

opdays <- operationaldays$daysop
```


