---
title: "New Analysis with Revisions"
author: "Elise Boos, Kelly Bruno"
date: "2023-11-10"
output: pdf_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
require(tidyverse)
#require(mapview)
require(sf)
require(raster)

require(snowfall)

#install.packages("rjags")
require(rjags)
library(camtrapR)
```

## Main critque that is addressed in new analysis:
* Cameras too close to one another, breaking independence assumption - *Gridding attempts to solve*
* Logged and Logged sites intermixed, breaking independence assumption - *Gridding attempts to solve*
* Lumped years frowned upon - *separated the years*
* single species is typically not done in these types of analyses - *switched to multi-species community model*


All of these have either more than 40 detections every year (2017-2019) or have over forty for at least one year after reviewing numbers for the other years and them being not too much lower than forty detections.

Species list = Agouti, Baird's Tapir, Crested Guan, Common Opossum, Gray Fox, Gray-headed Dove, Great Curassow, Jaguar, Ocellated Turkey, Ocelot, Paca, Puma, Red Brocket Deer, White-nosed Coati, White-tailed Deer

15 total species 

# Run it back

New analysis where years are separated - running single-season, multi-species models using JAGS. 2014, 2015, and 2016 were cut because they didn't have a good balance of logged and unlogged sites.  

Referenced code from: https://cran.r-project.org/web/packages/camtrapR/vignettes/camtrapr5.html

```{r load data, include=FALSE}
#Load and Clean Data 

##load dfs
load("dat.RData")
load("camop.RData")
covs_all <- read.csv("covs_fin.csv",stringsAsFactors = TRUE)

#fix logging column
covs_all$Logging <- ifelse(covs_all$Logging %in% c("No", "No Logging"), "No", "Yes")

#another option to rewrite factor
#library(forcats)
#covs_all2$Logging_fix <- fct_collapse(covs_all2$Logging, No = c("No","No Logging"), Yes = c("Yes","Logging"))

#missing lat and long 
covs_all[covs_all$site =="908F",]$Lat <- 17.66851300
covs_all[covs_all$site =="908F",]$Long <- -89.01283000



```

```{r fix easting northing, include=FALSE}
# these sites missing easting northing 
# 508H <- 288991 , 1929840
# 502C <- 292758, 1955010
# 503C <- 291102, 1950731
# 506C <- 291308 , 1950725
# 511C <- 282077, 1935175
# 506F <- 283678, 1935452

camop[camop$site == "508H",]$Easting <- 288991
camop[camop$site == "508H",]$Northing <- 1929840

camop[camop$site == "502C",]$Easting <- 292758
camop[camop$site == "502C",]$Northing <- 1955010

camop[camop$site == "503C",]$Easting <- 291102
camop[camop$site == "503C",]$Northing <- 1950731

camop[camop$site == "506C",]$Easting <- 291308
camop[camop$site == "506C",]$Northing <- 1950725

camop[camop$site == "511C",]$Easting <- 282077
camop[camop$site == "511C",]$Northing <- 1935175

camop[camop$site == "506F",]$Easting <- 283678
camop[camop$site == "506F",]$Northing <- 1935452

#turn into spatial feautre
cams.sf <- camop %>%
 # drop_na(Easting, Northing) %>%
  st_as_sf(coords = c('Easting','Northing'),
           crs=26716)

#getting logging values attatched to cams 
join <- left_join(cams.sf, covs_all, by = "site")
```

```{r mapview intro, eval=FALSE, include=FALSE}
#visualize
# mapview(join, zcol = "Logging")
# mapview(filter(join, year(Date.Placement) == 2019), zcol = "Logging") 
```

```{r making grid, include=FALSE}
### making grid
# Create an sf object for the grid
gridx <- seq(from = -89.1,to = -88.8, by = 0.01)
gridy <- seq(from = 17.4,to = 17.8, by = 0.01)


foo <- raster(xmn = -89.1, xmx = -88.8,
              ymn = 17.4, ymx =17.8,
              nrows = length(gridx),
              ncols = length(gridy)) %>% 
       rasterToPolygons() %>% 
       st_as_sf(
           crs=26716) %>% 
       mutate(group = 1:(length(gridx)*length(gridy))) %>% 
       st_cast("MULTIPOLYGON")

#transform to same crs
foo_trans <- st_transform(foo, crs(join))
#Now count how many data points exist in each grid
#mutate(foo,
 #      count = lengths(st_intersects(x = foo,  y = join))) -> foo
```

## Gridding

Making a 0.01 by 0.01 decimal degree grid (equating each grid to be a little bigger than 1km2). *"Our general recommendation is to separate cameras by as much as you can, with 500-1000m spacing being best, and anything >200m probably being acceptable."* These 1km2 grids allow for better distance separation between individual cameras by lumping those nearby into the same sampling grid. We are still able to reference the specific camera to account for differences in number of cameras in each grid.  


All cameras from every year in the grid colored by logged or not:

```{r cams and logging grid, echo=FALSE}
# Draw a graphic

ggplot() +
  geom_sf(data = foo)+
  geom_sf(data = join, aes(col=Logging))+
  xlim(-89.08, -88.9)+
  ylim(17.4, 17.71)
```

2018 cameras in the grid colored by logged or not:
```{r cams and logging for 2018, echo=FALSE}


ggplot() +
  geom_sf(data = foo)+
  geom_sf(data = filter(join, year(Date.Placement) == 2020), aes(col = Logging))+
  xlim(-89.05, -88.9)+
  ylim(17.48, 17.71)

#based on distribution of unlogged in 2020 and its proximity to logged sites I think we should drop it 
```

```{r mapview, eval=FALSE, include=FALSE}
# mapview(foo_trans)+
#   mapview(join, zcol = "Logging")
# 
# mapview(foo)+
# mapview(filter(join, year(Date.Placement) == 2018), zcol = "Logging")
```

```{r grid for everything, include=FALSE}
grid <- st_join(join,
                foo_trans,
                left = F)

mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

```

```{r}
belize_dem <- raster("./belize_dem_crop.tif")
plot(belize_dem)
plot(foo_trans$geometry, bg="transparent", add=TRUE)

ex <- extract(belize_dem, foo_trans, method = 'bilinear', fun = mean, na.rm = TRUE, df =TRUE)
summary(ex$belize_dem_crop)
```

# Draft Analysis By Year 

Need to increase iterations and chains for models, but preliminary code and workflow.

Used the threshold of 40 detection, some species did not meet that threshold for all 4 years, treated each year separately and outputted results for any species meeting that threshold in the single single year. 

Each call defines a model with a fixed effect of logging on occupancy probability of all species (have the option to add a species-specific (random) effect on occupancy probability). Furthermore, we specified effort as a covariate on detection probability (constant across species), and there are random effects of species on the intercepts of detection and occupancy probabilities.

All these test models only have logging as a covariate of interest. We are not sure if or what covariates we will ad in as fixed or random effects on occupancy and/or detection. 

```{r grid for 2017, include=FALSE}
grid_2017 <-filter(grid, year(Date.Placement) == 2017)

Logging_Grid_2017 <- grid_2017 %>%
  group_by(group) %>%
  summarise(Logging_Grid = mode(Logging))

#mapview(Logging_Grid_2017, zcol = "Logging_Grid")

Logging_Grid_2017.df <-dplyr::select(as.data.frame(Logging_Grid_2017), -geometry)

camgrid2017 <- left_join(grid_2017, Logging_Grid_2017.df, by = "group")

```

```{r grid for 2018, include=FALSE}
grid_2018 <-filter(grid, year(Date.Placement) == 2018)


Logging_Grid_2018 <- grid_2018 %>%
  group_by(group) %>%
  summarise(Logging_Grid = mode(Logging))

#mapview(Logging_Grid_2018, zcol = "Logging_Grid")

Logging_Grid_2018.df <-dplyr::select(as.data.frame(Logging_Grid_2018), -geometry)

camgrid2018 <- left_join(grid_2018, Logging_Grid_2018.df, by = "group")

```

```{r grid for 2019, include=FALSE}
grid_2019  <-filter(grid, year(Date.Placement) == 2019)

grid_spillover <- filter(grid, site %in% c("909B", "909C"))
grid_2019 <-rbind(grid_2019, grid_spillover)

Logging_Grid_2019 <- grid_2019 %>%
  group_by(group) %>%
  summarise(Logging_Grid = mode(Logging))

summary(Logging_Grid_2019$Logging_Grid)

#mapview(Logging_Grid_2019, zcol = "Logging_Grid")

Logging_Grid_2019.df <-dplyr::select(as.data.frame(Logging_Grid_2019), -geometry)

camgrid2019 <- left_join(grid_2019, Logging_Grid_2019.df, by = "group")


```

```{r grid for 2020, include=FALSE}
grid_2020 <-filter(grid, year(Date.Placement) == 2020)
# 2 2019 sites in here - move to 2019 or keep in 2020? Date placement was 2020

Logging_Grid_2020 <- grid_2020 %>%
  group_by(group) %>%
  summarise(Logging_Grid = mode(Logging))

#mapview(Logging_Grid_2020, zcol = "Logging_Grid")

Logging_Grid_2020.df <-dplyr::select(as.data.frame(Logging_Grid_2020), -geometry)

camgrid2020 <- left_join(grid_2020, Logging_Grid_2020.df, by = "group")

```


```{r 2017 cam grid, include=FALSE}

camgrid2017.df <- dplyr::select(as.data.frame(camgrid2017), -geometry)
cam_op <- cameraOperation(camgrid2017.df, stationCol = "group", cameraCol = "site", setupCol = "Date.Placement", byCamera = FALSE, allCamsOn = FALSE, camerasIndependent = FALSE,
                             retrievalCol = "Last.record", occasionStartTime = 0 , 
                             dateFormat = "%Y-%m-%d", writecsv = FALSE)


dat_2017 <-filter(dat, Year == 2017)
dat_2018 <-filter(dat, Year == 2018)
spillover_2017 <- dat_2018[1:17,]

dat_2017 <- rbind(dat_2017, spillover_2017)

datgrid2017 <- left_join(dat_2017, camgrid2017.df, by = "site")
```

```{r}
detectionMaps(CTtable = camgrid2017.df, recordTable = datgrid20172, Xcol = "Long", Ycol = "Lat", stationCol = "group", speciesCol = "Species", richnessPlot = T, speciesPlots = T, printLabels = T, plotR = T)

```

```{r test code, eval=FALSE, include=FALSE}

### TEST CODE ###
detect_hist_agouti <- detectionHistory(recordTable = datgrid2017, species = "Agouti",
                                       camOp = cam_op, output = "binary", stationCol = "group", 
                                       speciesCol = "Species", day1 ="station", 
                                       recordDateTimeCol = "date_time_obs", 
                                       recordDateTimeFormat =  "%Y-%m-%d %H.%M", 
                                       timeZone = "UTC", occasionLength = 1, 
                                       includeEffort = TRUE, scaleEffort = FALSE, 
                                       writecsv = FALSE)

detect_hist_puma <- detectionHistory(recordTable = datgrid2017, species = "Puma",
                                       camOp = cam_op, output = "binary", stationCol = "group", 
                                       speciesCol = "Species", day1 ="station", 
                                       recordDateTimeCol = "date_time_obs", 
                                       recordDateTimeFormat =  "%Y-%m-%d %H.%M", 
                                       timeZone = "UTC", occasionLength = 1, 
                                       includeEffort = TRUE, scaleEffort = FALSE, 
                                       writecsv = FALSE)
ylist <- list(Agouti=detect_hist_agouti$detection_history, Puma=detect_hist_puma$detection_history)

obs <- list(effort = detect_hist_agouti$effort)


Logging_Grid_2017.df$Logging_Grid <- as.factor(Logging_Grid_2017.df$Logging_Grid)
data_list <- list(ylist    = ylist,
                   siteCovs = Logging_Grid_2017.df,
                   obsCovs  = obs)  # is identical for all species

 modelfile1 <- tempfile(fileext = ".txt")
mod.jags <- communityModel(data_list,
                            occuCovs = list(fixed = "Logging_Grid"),
                            detCovsObservation = list(fixed = "effort"),
                            intercepts = list(det = "ranef", occu = "ranef"),
                            modelFile = modelfile1)

summary(mod.jags)


fit.jags <- fit(mod.jags,
                 n.iter = 1000,
                 n.burnin = 500,
                 chains = 3)

fit_summary <- summary(fit.jags)

DT::datatable(round(fit_summary$statistics, 3))

 plot_effects(mod.jags,
              fit.jags,
              submodel = "state")
 
  plot_effects(mod.jags,
              fit.jags,
              submodel = "det")
  

```

```{r 2017 species, include=FALSE}
Logging_Grid_2017.df$Logging_Grid <- as.factor(Logging_Grid_2017.df$Logging_Grid)

Logging_Grid_2017.df <- left_join(Logging_Grid_2017.df, ex, by = c("group" = "ID"))

 # list of detection histories
datgrid2017$Species <- as.character(datgrid2017$Species)
#remove error rows if throws error 
datgrid20172 <- datgrid2017[-52,]
#check if runs all species

#get count of species
# count <- table(datgrid20172$Species)
# count
# #count
# list <- subset(count, count > 40)
# 
# species_list <- names(list)

species_list <- c("Agouti", "Baird's Tapir", "Crested Guan", "Common Opossum", "Gray Fox", "Gray-headed Dove", "Great Curassow", "Jaguar", "Ocellated Turkey", "Ocelot", "Paca", "Puma", "Red Brocket Deer", "White-nosed Coati", "White-tailed Deer")
species_list

```

## 2017 Prelim Results

```{r run 2017 model, include=FALSE}
 DetHist_list <- lapply(species_list, FUN = function(x) {
   detectionHistory(recordTable = datgrid20172, species = x,
                                       camOp = cam_op, output = "binary", stationCol = "group", 
                                       speciesCol = "Species", day1 ="station", 
                                       recordDateTimeCol = "date_time_obs", 
                                       recordDateTimeFormat =  "%Y-%m-%d %H.%M", 
                                       timeZone = "UTC", occasionLength = 1, 
                                       includeEffort = TRUE, scaleEffort = FALSE, 
                                       writecsv = FALSE)}
 )

 # assign species names to the list items
 names(DetHist_list) <- species_list
 #double check correct naming

 # note, DetHist_list is a list containing a list for each species
 ylist <- lapply(DetHist_list, FUN = function(x) x$detection_history)

data_list <- list(ylist    = ylist,
                   siteCovs = Logging_Grid_2017.df,
                   obsCovs  = list(effort = DetHist_list[[1]]$effort)) 

# text file to save the model
 modelfile1 <- tempfile(fileext = ".txt")
```

```{r model code 2017, echo=TRUE}
 mod.jags <- communityModel(data_list,
                            occuCovs = list(fixed = "Logging_Grid", ranef = "belize_dem_crop"),
                            detCovsObservation = list(fixed = "effort"),
                            intercepts = list(det = "ranef", occu = "ranef"),
                            modelFile = modelfile1)
summary(mod.jags)

#will need to increase iterations and maybe chains when running for real
 fit.jags <- fit(mod.jags,
                 n.iter = 10000,
                 n.burnin = 500,
                 chains = 4)
``` 

```{r model figures 2017, echo=FALSE}
 fit_summary <- summary(fit.jags)
 as.data.frame(fit_summary$statistics)
 
 #DT::datatable(round(fit_summary$statistics, 3))
 
  plot_effects(mod.jags,
              fit.jags,
              submodel = "state")
  
   plot_effects(mod.jags,
              fit.jags,
              submodel = "det")
   
    plot_coef(mod.jags,
           fit.jags,
           submodel = "state")
   
```

```{r 2018 cam grid, include=FALSE}
camgrid2018.df <- dplyr::select(as.data.frame(camgrid2018), -geometry)
cam_op <- cameraOperation(camgrid2018.df, stationCol = "group", cameraCol = "site", setupCol = "Date.Placement", byCamera = FALSE, allCamsOn = FALSE, camerasIndependent = FALSE,
                             retrievalCol = "Last.record", occasionStartTime = 0 , 
                             dateFormat = "%Y-%m-%d", writecsv = FALSE)

#some of the 2017 data goes through 2018 !!!
dat_2018  <-filter(dat, Year == 2018)
#removing 2017 data add back into 2017 or into this tbd 
dat_2018 <- dat_2018[-(1:17),]
datgrid2018 <- left_join(dat_2018, camgrid2018.df, by = "site")


 # list of detection histories
datgrid2018$Species <- as.character(datgrid2018$Species)
#remove error rows 
#datgrid20172 <- datgrid2017[-52,]



#get count of species
# count <- table(datgrid2018$Species)
# count
# #count
# list <- subset(count, count > 40)
# 
# species_list <- names(list)
# species_list
```

## 2018 Prelim Results

```{r run 2018 code, include=FALSE}

 DetHist_list <- lapply(species_list, FUN = function(x) {
   detectionHistory(recordTable = datgrid2018, species = x,
                                       camOp = cam_op, output = "binary", stationCol = "group", 
                                       speciesCol = "Species", day1 ="station", 
                                       recordDateTimeCol = "date_time_obs", 
                                       recordDateTimeFormat =  "%Y-%m-%d %H.%M", 
                                       timeZone = "UTC", occasionLength = 1, 
                                       includeEffort = TRUE, scaleEffort = FALSE, 
                                       writecsv = FALSE)}
 )

 # assign species names to the list items
 names(DetHist_list) <- species_list

 # note, DetHist_list is a list containing a list for each species
 ylist <- lapply(DetHist_list, FUN = function(x) x$detection_history)

 Logging_Grid_2018.df$Logging_Grid <- as.factor(Logging_Grid_2018.df$Logging_Grid)
 Logging_Grid_2018.df <- left_join(Logging_Grid_2018.df, ex, by = c("group" = "ID"))
 
data_list <- list(ylist    = ylist,
                   siteCovs = Logging_Grid_2018.df,
                   obsCovs  = list(effort = DetHist_list[[1]]$effort)) 

# text file to save the model
 modelfile2 <- tempfile(fileext = ".txt")

```

```{r model code 2018, echo=TRUE}
 mod.jags <- communityModel(data_list,
                            occuCovs = list(fixed = "Logging_Grid", ranef = "belize_dem_crop"),
                            detCovsObservation = list(fixed = "effort"),
                            intercepts = list(det = "ranef", occu = "ranef"),
                            modelFile = modelfile2)
summary(mod.jags)

 fit.jags <- fit(mod.jags,
                 n.iter = 10000,
                 n.burnin = 500,
                 chains = 3)
```

```{r model graphs 2018, echo=FALSE}
 
 fit_summary <- summary(fit.jags)
 as.data.frame(fit_summary$statistics)
 #DT::datatable(round(fit_summary$statistics, 3))
 
  plot_effects(mod.jags,
              fit.jags,
              submodel = "state")
  
   plot_effects(mod.jags,
              fit.jags,
              submodel = "det")
   
    plot_coef(mod.jags,
           fit.jags,
           submodel = "state")

```


```{r 2019 camgrid, include=FALSE}
camgrid2019.df <- dplyr::select(as.data.frame(camgrid2019), -geometry)
cam_op <- cameraOperation(camgrid2019.df, stationCol = "group", cameraCol = "site", setupCol = "Date.Placement", byCamera = FALSE, allCamsOn = FALSE, camerasIndependent = FALSE,
                             retrievalCol = "Last.record", occasionStartTime = 0 , 
                             dateFormat = "%Y-%m-%d", writecsv = FALSE)


dat_2019  <-filter(dat, Year == 2019)

#add in spill over from 2020
dat_2020 <-filter(dat, Year == 2020)
spillover_2020 <- dat_2020[1:50,]

dat_2019 <- rbind(dat_2019, spillover_2020)


datgrid2019 <- left_join(dat_2019, camgrid2019.df, by = "site")


 # list of detection histories
datgrid2019$Species <- as.character(datgrid2019$Species)
#remove error rows 
datgrid20192 <- datgrid2019[-2227,]

#get count of species
count <- table(datgrid2019$Species)
count
#count
list <- subset(count, count > 40)

species_list <- names(list)
species_list

```

## 2019 Prelim Results

```{r results 2019, include=FALSE}
 DetHist_list <- lapply(species_list, FUN = function(x) {
   detectionHistory(recordTable = datgrid20192, species = x,
                                       camOp = cam_op, output = "binary", stationCol = "group", 
                                       speciesCol = "Species", day1 ="station", 
                                       recordDateTimeCol = "date_time_obs", 
                                       recordDateTimeFormat =  "%Y-%m-%d %H.%M", 
                                       timeZone = "UTC", occasionLength = 1, 
                                       includeEffort = TRUE, scaleEffort = FALSE, 
                                       writecsv = FALSE)}
 )

 # assign species names to the list items
 names(DetHist_list) <- species_list

 # note, DetHist_list is a list containing a list for each species
 ylist <- lapply(DetHist_list, FUN = function(x) x$detection_history)

 Logging_Grid_2019.df$Logging_Grid <- as.factor(Logging_Grid_2019.df$Logging_Grid)
 
data_list <- list(ylist    = ylist,
                   siteCovs = Logging_Grid_2019.df,
                   obsCovs  = list(effort = DetHist_list[[1]]$effort)) 

# text file to save the model
 modelfile1 <- tempfile(fileext = ".txt")

```

```{r model run for 2019, echo=TRUE}
 mod.jags <- communityModel(data_list,
                            occuCovs = list(fixed = "Logging_Grid"),
                            detCovsObservation = list(fixed = "effort"),
                            intercepts = list(det = "ranef", occu = "ranef"),
                            modelFile = modelfile1)
summary(mod.jags)

 fit.jags <- fit(mod.jags,
                 n.iter = 1000,
                 n.burnin = 500,
                 chains = 3)
```

```{r model graphs for 2019, echo=FALSE}
 fit_summary <- summary(fit.jags)
 as.data.frame(fit_summary$statistics)
 
 #DT::datatable(round(fit_summary$statistics, 3))
 
  plot_effects(mod.jags,
              fit.jags,
              submodel = "state")
  
   plot_effects(mod.jags,
              fit.jags,
              submodel = "det")
   
    plot_coef(mod.jags,
           fit.jags,
           submodel = "state")

```


```{r camgrid 2020, include=FALSE}
camgrid2020.df <- dplyr::select(as.data.frame(camgrid2020), -geometry)
cam_op <- cameraOperation(camgrid2020.df, stationCol = "group", cameraCol = "site", setupCol = "Date.Placement", byCamera = FALSE, allCamsOn = FALSE, camerasIndependent = FALSE,
                             retrievalCol = "Last.record", occasionStartTime = 0 , 
                             dateFormat = "%Y-%m-%d", writecsv = FALSE)


dat_2020  <-filter(dat, Year == 2020)
# #removing 2017 data add back into 2017 or into this tbd 
 dat_2020 <- dat_2020[-(1:50),]
datgrid2020 <- left_join(dat_2020, camgrid2020.df, by = "site")


 # list of detection histories
datgrid2020$Species <- as.character(datgrid2020$Species)
#remove error rows 
datgrid20202 <- datgrid2020[-1019,]

#get count of species
count <- table(datgrid20202$Species)
count
#count
list <- subset(count, count > 40)

species_list <- names(list)
species_list

```

## 2020 Prelim Results

```{r 2020 results, include=FALSE}
 DetHist_list <- lapply(species_list, FUN = function(x) {
   detectionHistory(recordTable = datgrid20202, species = x,
                                       camOp = cam_op, output = "binary", stationCol = "group", 
                                       speciesCol = "Species", day1 ="station", 
                                       recordDateTimeCol = "date_time_obs", 
                                       recordDateTimeFormat =  "%Y-%m-%d %H.%M", 
                                       timeZone = "UTC", occasionLength = 1, 
                                       includeEffort = TRUE, scaleEffort = FALSE, 
                                       writecsv = FALSE)}
 )

 # assign species names to the list items
 names(DetHist_list) <- species_list

 # note, DetHist_list is a list containing a list for each species
 ylist <- lapply(DetHist_list, FUN = function(x) x$detection_history)

 Logging_Grid_2020.df$Logging_Grid <- as.factor(Logging_Grid_2020.df$Logging_Grid)
 
data_list <- list(ylist    = ylist,
                   siteCovs = Logging_Grid_2020.df,
                   obsCovs  = list(effort = DetHist_list[[1]]$effort)) 

# text file to save the model
 modelfile1 <- tempfile(fileext = ".txt")

```

```{r model for 2020, echo=TRUE}
 mod.jags <- communityModel(data_list,
                            occuCovs = list(fixed = "Logging_Grid"),
                            detCovsObservation = list(fixed = "effort"),
                            intercepts = list(det = "ranef", occu = "ranef"),
                            modelFile = modelfile1)
summary(mod.jags)

 fit.jags <- fit(mod.jags,
                 n.iter = 1000,
                 n.burnin = 500,
                 chains = 3)
```

```{r model figures for 2020, echo=FALSE}
 fit_summary <- summary(fit.jags)
 as.data.frame(fit_summary$statistics)
 
 #DT::datatable(round(fit_summary$statistics, 3))
 
  plot_effects(mod.jags,
              fit.jags,
              submodel = "state")
  
   plot_effects(mod.jags,
              fit.jags,
              submodel = "det")
   
    plot_coef(mod.jags,
           fit.jags,
           submodel = "state")

```

