---
title: "Code example 2017"
author: "Kelly Bruno"
date: '2022-10-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require("knitr")
opts_knit$set(root.dir = "~/Belize-MP-Bruno-Boos")
getwd()
```

```{r load packages etc}
getwd()
#packages
require(tidyverse)
require(dplyr)
require(sf)
require(leaflet)
require(mapview)
require(readxl)
require(lubridate)
require(readxl)

library(unmarked)
library(AICcmodavg)
#make mapview work on a mac
mapviewOptions(fgb = FALSE)

library(camtrapR)

```

```{r load data}

#load data
records_2017 <- read.csv("./By_year/2017_Records_Yalbac_data.csv", stringsAsFactors = T)
```

CLEANING

```{r cleaning 2017}
records_2017_clean <- records_2017 %>%
  mutate(site = paste0(Cam.nr,series)) %>%
  drop_na(Cam.nr)

unique(records_2017_clean$site)

cameraops2017 <- read.csv("./2017/Cameras_2017.csv", header = T)

#issue with colname import
colnames(cameraops2017)[colnames(cameraops2017)=="ï..Name"] <- "Name"

unique(cameraops2017$Name)

datafor2017 <- full_join(records_2017_clean, cameraops2017, by = c("site" = "Name"))

#set as date 
datafor2017$Date.Placement <- as.Date(datafor2017$Date.Placement, "%d-%b-%y")
datafor2017$Date.Removed <- as.Date(datafor2017$Date.Removed, "%d-%b-%y")

#calc end date for cam
datafor2017_edit <- datafor2017 %>%
  mutate(Last.Record = Date.Placement + Days + 1)

#fix obs date for 10260313 in 710E
datafor2017_edit[1248, ]$Day <- 26

#visit day
datafor2017_visit <- datafor2017_edit %>%
  mutate(obsdate = as.Date(paste0(Year,"-",Month,"-",Day), "%Y-%m-%d")) %>%
  mutate(visit = as.numeric((obsdate - Date.Placement) + 1))

#error finder 
datafor2017_errors <- datafor2017_visit %>%
  mutate(errors = ifelse(obsdate > Last.Record, TRUE, FALSE)) %>%
  mutate(opp_errors = ifelse(obsdate < Date.Placement, TRUE, FALSE)) 
summary(datafor2017_errors$errors)
summary(datafor2017_errors$opp_errors)

#fix errors 

#last record update for after typos fixed
fixerrors_2017 <- datafor2017_errors %>%
  mutate(max.date = if_else(as.Date(obsdate) > as.Date(Last.Record), obsdate, Last.Record)) %>%
  group_by(site) %>%
  summarize(Last.Record_clean = max(max.date)) 
  
datafor2017_int <- full_join(datafor2017_errors, fixerrors_2017, by = "site")

fixerrors_2017_cont <- datafor2017_int %>%
  mutate(min.date = if_else(as.Date(obsdate) < as.Date(Date.Placement), obsdate, Date.Placement)) %>%
  group_by(site) %>%
  summarize(First_Obs = min(min.date))

datafor2017_final <- full_join(datafor2017_int, fixerrors_2017_cont, by = "site")

#check final
datafor2017_check <- datafor2017_final %>%
  mutate(errors_final = ifelse(obsdate > Last.Record_clean, TRUE, FALSE)) %>%
  mutate(errors_final2 = ifelse(obsdate < First_Obs, TRUE, FALSE))

data2017_final <- datafor2017_final  %>%
  mutate(visit = as.numeric((obsdate - First_Obs + 1))) %>%
  select(site, 1:14, 17,18,34,19:21,25:33, First_Obs, Last.Record_clean) %>%
  mutate(date_time_obs = paste(obsdate, sprintf("%.2f", round(Hour.Exact, 2)))) %>%
  drop_na(Hour.Exact)
```

2017 Occupancy Frame

```{r making occupancy dataframe 2017}

require(camtrapR)
camdataprep_2017 <- data2017_final %>%
  select(site, First_Obs, Days, Last.Record_clean, Northing, Easting)
camopdataframe2017 <- camdataprep_2017[!duplicated(camdataprep_2017$site), ]

camop2017 <- cameraOperation(camopdataframe2017, stationCol = "site", setupCol = "First_Obs", retrievalCol = "Last.Record_clean", occasionStartTime = 0 , dateFormat = "%Y-%m-%d", writecsv = FALSE)


detect_hist_2017_grayfox <- detectionHistory(recordTable = data2017_final, species = "Gray fox", camOp= camop2017, output = "binary", stationCol = "site", speciesCol = "species.name", day1 ="station", recordDateTimeCol = "date_time_obs", recordDateTimeFormat =  "%Y-%m-%d %H.%M", timeZone = "UTC", occasionLength = 1, includeEffort = TRUE, scaleEffort = FALSE, writecsv = FALSE)

detectionhist2017 <- detect_hist_2017_grayfox$detection_history

```

Start modeling

```{r model test}
covariates <- read_xlsx("Covariates_cams.xlsx")

cams700 <- covariates[188:263,]

dataframe_det <- as.data.frame(detectionhist2017)
dataframe_det$rn <- row.names(dataframe_det)


join <- left_join(dataframe_det,cams700, by = c("rn" = "CameraID"))
cams700_clean <- select(join, 116:215)

#cov_list <- as.list(cams700_clean) #convert covariate frame into list
#df_cov <- lapply(cov_list, as.data.frame) #convert to list of dfs

#check all covs - for some reason some have been converted to factors?
cams700_clean$Avg_pr_2017 <- as.numeric(cams700_clean$Avg_pr_2017) #for this test I am only fixing avg. precip 2017 but will fix other errors later

unmarkedFrame <- unmarkedFrameOccu(detectionhist2017, siteCovs = cams700_clean)

#standardizing variables - need to do others but this is a test
unmarkedFrame@siteCovs$Avg_pr_2017 <- scale(unmarkedFrame@siteCovs$Avg_pr_2017)

unmarkedFrame@siteCovs$Canopy_Height_m <- scale(unmarkedFrame@siteCovs$Canopy_Height_m)

summary(unmarkedFrame)

```

Model Time

```{r ooo model}

modlist <- list()

modlist[["intercept"]] <- fm1 <- occu(formula = ~1 # detection formula first
                     ~1,# occupancy formula second,
           data = unmarkedFrame)

##############################################
###To get real estimate of occupancy (with 95% CI)
predict(fm, 
        newdata = data.frame(site = 1),
        type = "state")
# ~0.59

# To get real estimate of detection (with 95% CI)
predict(fm, 
        newdata = data.frame(site = 1),
        type = "det")
# ~0.13
##############################################

modlist[["precip"]] <- fm2 <- occu(formula = ~1
                     ~Avg_pr_2017,
           data = unmarkedFrame)

modlist[["canopy"]] <- fm3 <- occu(formula = ~1
                                   ~Canopy_Height_m,
                                   data = unmarkedFrame)

##add one with both covs

aictab(modlist)

```

Reminder: There are two types of covariates:

    Observation-level covariates, which are different for every combination of site and survey (e.g. search effort per visit per site). Each one of these variables should be included as an element of a list, and each element of the list should have a row for every site and a column for every survey. In our example, that would be 100 rows and 10 columns.

    Site-level covariates do not change through time, but differ between each site. For example, measures of habitat type or other landscape metrics like the straight-line distance to a road. Site covariates should be included in a dataframe with a row for every site and a column for each site variable.
