---
title: "Unmarked tutorial"
author: "Kelly Bruno"
date: '2022-10-03'
output: html_document
---

Occupancy models estimate where species occur (a biological process) while accounting for imperfect detection (an observation process). This is important because the probability of detecting a species during any survey is almost always substantially below 1. Without accounting for imperfect detection, estimates of species distribution are considerably biased.

A basic occupancy model estimates two parameters:

detection probability (pi,t, the probability of finding a species at time t if it is present at site i)
occupancy (Ψi, the probability that a species occurs at site i).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("~/Belize-MP-Bruno-Boos/occupancyworkshop-master/occupancyworkshop-master")


#load detection history (100 sites with 10 visits each)
detection_history <- read.csv("detection_history.csv", 
                              # First variable ("X") has row.names but not data
                              row.names = "X") 

# Examine data
head(detection_history)

```

The basic single-season form assumes that occupancy status (presence or absence) does not change during the study.

```{r data}
library(unmarked)

sample.unmarkedFrame_simple <- unmarkedFrameOccu( # y is a matrix with observed detection history 
                                          # (0's and 1's, one row per site, one column per survey)
                                      y = as.matrix(detection_history)) 

# S4 class for occupancy model data
summary(sample.unmarkedFrame_simple)

```






